---
title: "Zillow Exploratory Data Analysis KEY"
author: "John Adams"
output: github_document
---
#Zillow Exploratory Data Analysis Worksheet

```{r load the tidyverse and skimr libraries, warning=FALSE}
library(tidyverse)
library(skimr)
library(readr)
```

#Download the Data

```{r Download the data}
zillow <- read_csv("https://raw.githubusercontent.com/jwadams25/LearningR/master/housingdata.csv")

```
## Load The Data
```{r load the data}
zillow <- read.csv("file:///Users/j.adams/projects/LearningR/housingdata.csv", stringsAsFactors=FALSE)
```

## Exploring the Data

Use this code to help answer questions in the Exploring the Data section of your worksheet

```{r viewing the data}

head(zillow)

```


# Visualize the Data

To begin, let's explore some distributions of both price and living area. As you have learned, a historgram is a great option. 

## Create a histogram showing the distribution of Price

```{r Price histogram}

p_hist <- zillow %>%
  ggplot(aes(Price)) +
  geom_histogram(binwidth = 50000, fill = "navy", color = "black") +
  labs(title = "Distribution of House Prices", x = "Price", y = "Number of Houses") +
  scale_x_continuous(limits = c(0, 800000),
                   breaks = seq(0, 800000, 50000))
p_hist

```

## Create a histogram showing the distribution of Living Area

```{r Living Area histogram}
la_hist <- zillow %>%
          ggplot(aes(Living.Area)) +
  geom_histogram(binwidth = 200, fill = "red", color = "black") +
  labs(title =" Distribution of Living Areas", x = "Living Area (sq-ft)", 
       y = "Number of Houses") +
  scale_x_continuous(limits = c(0, 5600),
                     breaks = seq(0, 5600, 200)) +
  scale_y_continuous(limits = c(0, 300),
                     breaks = seq(0, 300, 50))

la_hist

```

# Summary statistics for histograms 

Let's now look at the summary statistics for the quantitative variables, Price and living area, displayed in the histograms.

##Summary Statistics for Price
```{r Price Summary Stats, warning=FALSE}

ssprice <- zillow %>% 
          summarize(avg = mean(Price), med = median(Price), standard_dev = sd(Price), 
                    iqr = IQR(Price))
ssprice

```

##Summary Statistics for Living Area
```{r Living Area Summary Stats, warning=FALSE}
ssarea <- zillow %>% 
          summarize(avg = mean(Living.Area), med = median(Living.Area), 
                    standard_dev = sd(Living.Area), 
                    iqr = IQR(Living.Area))
ssarea

```

##Summary Statistics for Price and Living Area 
Here is another way to calculate summary statistics for multiple variables at the same time. 
```{r price and living area summary stats}

ss_area_price <- zillow %>%
                  select(Price, Living.Area) %>%
                  skim()
ss_area_price

```

# Relationship Between Price and Living Area

##Create a Scatter Plot

Now let's see if there is an association between the price and living area. Create a scatter plot of Price vs Living Area.
```{r scatter plot}

pvsa <- zillow %>%
          ggplot(aes(x= Living.Area, y = Price)) +
          geom_point(alpha = 0.25) +
          stat_smooth(method = 'lm') +
          labs(title ="Price vs. Living Area", 
               x = "Living Area (sq. ft)", y = "Price") +
          scale_y_continuous(limits = c(0, 800000),
                     breaks = seq(0, 800000, 250000))
pvsa
```

## Creating the Model

Let's now build a linear model to gain more insight into the relationship between Prices and Living Area.

```{r create the model}
zillowmodel <- lm(Price ~ Living.Area, zillow)
zillowmodel
cor(zillow$Price, zillow$Living.Area)

```

# Conditional Distributions

Houses can use different types of heat. Let's see if there is different between the types of heat in newly constructed houses and not new constructed houses. 

```{r stacked bar and two way table}
heat_construction <-  zillow %>%
                    ggplot(aes(x = New.Construct, fill = Heat.Type)) +
                    geom_bar(position = "fill")
heat_construction
```

## Create a Two-Way Table

```{r two way table}
heattotals <- zillow %>%
            group_by(New.Construct, Heat.Type) %>%
            summarize(num = n()) %>%
            spread(New.Construct, num)
heattotals
```








