---
title: "Zillow Exploratory Data Analysis Worksheet"
author: "John Adams"
output: github_document
---
#Zillow Exploratory Data Analysis Worksheet

```{r load the tidyverse library}
library(tidyverse)
library(skimr)
```

# Exploring the Data

Use this code to help answer questions in the Exploring the Data section of your worksheet

```{r viewing the data}
#after you run this code be sure to click the arrow to see all variables
head(NAME_OF_DATASET)

```


# Visualize the Data

To begin, let's explore some distributions of both price and living area. As you have learned, a historgram is a great option. 

### Create a histogram showing the distribution of Price

```{r Price histogram}
p_hist <- NAME_OF_DATASET %>%
  ggplot(aes(QUANTITATIVE_VARIABLE)) +
  geom_histogram(binwidth = NUMBER, fill = "navy", color = "black") +
  labs(title = "Distribution of House Prices", x = "X-Axis Title", y = "Y-Axis Title") +
  scale_x_continuous(limits = c(0, 800000),
                   breaks = seq(0, 800000, 50000))
p_hist
```

## Create a histogram showing the distribution of Living Area

```{r Living Area Histogram}
la_hist <- NAME_OF_DATASET %>%
          ggplot(aes(QUANTITATIVE_VARIABLE)) +
  geom_histogram(binwidth = NUMBER, fill = "red", color = "black") +
  labs(title =" Distribution of Living Areas", x = "X-Axis Title", 
       y = "y-Axis Title") +
  scale_x_continuous(limits = c(0, 5600),
                     breaks = seq(0, 5600, 200)) +
  scale_y_continuous(limits = c(0, 300),
                     breaks = seq(0, 300, 50))
la_hist
```

# Summary statistics for histograms 

Let's now look at the summary statistics for the quantitative variables, Price and living area, displayed in the histograms.

##Summary Statistics for Price
```{r Price Summary Stats}

ssprice <- NAME_OF_DATASET %>% 
          summarize(avg = mean(QUANTITATIVE_VARIABLE), med = median(QUANTITATIVE_VARIABLE), 
                    standard_dev = sd(QUANTITATIVE_VARIABLE), 
                    iqr = IQR(QUANTITATIVE_VARIABLE))
ssprice
```

##Summary Statistics for Living Area
```{r Living Area Summary Statistics}
sslarea <- NAME_OF_DATASET %>% 
          summarize(avg = mean(QUANTITATIVE_VARIABLE), med = median(QUANTITATIVE_VARIABLE), 
                    standard_dev = sd(QUANTITATIVE_VARIABLE, 
                                      five_number = fivenum(QUANTITATIVE_VARIABLE)), 
                    iqr = IQR(QUANTITATIVE_VARIABLE))
sslarea
```

##Summary Statistics for Price and Living Area 
```{r price and living area summary stats}
ss_area_price <- NAME_OF_DATASET %>%
                  select(QUANTITATIVE_VARIABLE, QUANTITATIVE_VARIABLE) %>%
                  skim()
ss_area_price
```

# Relationship Between Price and Living Area

##Create a Scatter Plot

Now let's see if there is an association between the price and living area. Create a scatter plot of Price vs Living Area.

```{r scatter plot price vs living area}

pvsa <- NAME_OF_DATASET %>%
          ggplot(aes(x= IndependentVariable, y = DependentVariable)) +
          geom_point(alpha = 1) +
          stat_smooth(method = 'lm') +
          labs(title ="Main Title", x = "Independent Variable", y = "Dependent Variable") +
          scale_y_continuous(limits = c(0, 800000),
                     breaks = seq(0, 800000, 250000))

```

## Creating the Model

Let's now build a linear model to gain more insight into the relationship between Prices and Living Area.

```{r create the model}


NAME_YOUR_MODEL <- lm(DependentVariable~IndependentVariable, NAME_OF_DATASET)
NAME_YOUR_MODEL
cor(NAME_OF_DATASET$DependentVariable, NAME_OF_DATASET$IndependentVariable)

```
# Conditional Distributions

Houses can use different types of heat. Let's see if there is different between the types of heat in newly constructed houses and not new constructed houses. 

##Create a Segmented Bar Graph
```{r stacked bar and two way table}
heat_construction <-  NAME_OF_DATASET %>%
                    ggplot(aes(x = CATEGORICAL_VARIABLE, fill = CATEGORICAL_VARIABLE)) +
                    geom_bar(position = "fill")
heat_construction
```

## Create a Two-Way Table
```{r two way table}
heattotals <- NAME_OF_DATASET %>%
            group_by(New.Construct, CATEGORICAL_VARIABLE) %>%
            summarize(num = n()) %>%
            spread(New.Construct, num)
heattotals
```




